\chapter*{Conclusion}
\addcontentsline{toc}{chapter}{Conclusion}

Domain-specific planning has historically been neglected as not general enough
and, therefore, theoretically unimportant.
Our work shows that there are more than enough problems
left to solve when planning with prior domain knowledge,
both theoretically and in practical approaches.

To support this statement,
we have discussed the theoretical challenges underlying
planning in general and how they change when
domain knowledge is added.
Using variants of the Transport domain from the International Planning Competition,
we have shown an approach to analyzing,
planning, and evaluating domain-specific planners.
To make each step along the way easier, we have developer TransportEditor,
a transportation planning system, as an intuitive graphical desktop application,
which was recently accepted to the System Demonstrations and Exhibits
track at the 27$^\mt{th}$ International Conference on Automated Planning and Scheduling
in Pittsburgh, USA \citep{Skopek2017}. We have also developed a benchmarker project
for automatizing experimental evaluation of planners, which could be used for benchmarking planners on
other planning domains with just minor modifications.

To summarize our work, we have experimentally evaluated and discussed several approaches to
domain-specific planning on the Transport domain, both on the sequential and temporal
variants.
We have achieved comparable results to state-of-the-art domain-independent planners. Their performance is generally very impressive, given the difficulty of the problem they are solving. Despite the broad misconception that they are not useful in practice,
we have not managed to beat all of them even when leveraging domain-specific knowledge.

There remain more promising approaches to apply to the Transport domain
and domain-specific planning in general.
To conclude this thesis, we list a few, in our opinion, perspective methods for domain-specific planning, which were not
evaluated in this work due to time constraints:
\begin{itemize}
\item Hierarchical Task Networks (HTN) --- 
HTN planning \citet[Chapter~11]{Ghallab2004} uses \textit{tasks},
a higher level and usually domain-specific description of sequences of operators
to carry out some goal. An HTN planner decomposes these tasks and embeds them in a classical plan. This approach has been thoroughly studied and is arguably one of the most used in practice today.

\item Pointer Networks and Reinforcement Learning --- 
A recent attempt at training special architectures of neural networks to solve TSP problem instances using reinforcement learning by \citet{Bello2016} shows
reasonable promise for the future. While this technique is highly experimental at the moment,
neural networks have helped push the limits in other fields before.

\item Learning a domain-specific heuristic function --- Another neural network
approach aims to help solve a standard problem: coming up with a good heuristic for a domain. \citet{Chen2011}
train a neural network to use as a heuristic for state space search,
which may help when creating a heuristic is simply too challenging or time-consuming.
This approach has also been part of the recent DeepStack success at
beating human players in Poker \citep{Moravcik2017}, which is a good example of a problem with a large search space.
\end{itemize}